---
title: "Umesh Sai Teja Poola's Lab 4 Report"
author: "Umesh Sai Teja Poola"
date: "2024-12-09"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

```{r}
library(rvest)
library(dplyr)
library(ggplot2)
library(Lahman)

```


```{r}

# Define the URL for scraping
url <- "https://www.baseball-reference.com/awards/hof_2024.shtml"

# Read the HTML content
html <- read_html(url)

# Extract the tables
tables <- html_table(html)

# Select the first table (assuming it's the relevant one)
raw_data <- tables[[1]]


```


```{r}
# Assign column names from the first row, then remove the first row
colnames(raw_data) <- raw_data[1, ]
raw_data <- raw_data[-1, ]

# Select relevant columns and convert to the appropriate types
cleaned_data <- raw_data %>%
  select(Name, Votes = `Votes`) %>%
  mutate(
    Votes = as.numeric(Votes),
    yearID = 2024,
    votedBy = "BBWAA",
    ballots = 385,
    needed = 289,
    inducted = ifelse(Votes >= 289, "Y", "N"),
    category = NA,
    needed_note = NA
  )

```

```{r}
# Add the cleaned data to the existing HallOfFame dataset
combined_data <- bind_rows(
  HallOfFame,
  cleaned_data %>% rename(player_name = Name, votes = Votes)
)


```

```{r}
# Display the structure of the combined dataset
str(combined_data)


```

``` {r}
# Export the combined dataset
write_csv(combined_hof, "HallOfFame.csv", na = "")

```

```{r}
# Create a bar chart for inductees by year
ggplot(combined_data, aes(x = yearID, fill = inducted)) +
  geom_bar() +
  labs(
    title = "Baseball Hall of Fame Inductees (1936–2024)",
    x = "Year",
    y = "Count of Inductees"
  ) +
  theme_minimal()


```

## Reflection and Notes

- **Challenges Faced**: Matching player names with IDs required additional cleaning, as some names included special characters or unconventional formatting.
- **Future Improvements**: Automating the detection and handling of discrepancies in player names will improve accuracy and efficiency.
- **Key Takeaways**: Understanding how to structure web-scraped data to match a predefined schema and combining it with existing datasets for analysis.

